---
title: "Course 5 Week 2 Project"
output: html_document
---
```{r "setup", include = FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/Christopher/Desktop/Files Not Yet Backed-up/R Files/Course 5/RepData_PeerAssessment1")
```

```{r no sci_notattion}
inline_hook <- function(x) {
  if (is.numeric(x)) {
    format(x, digits = 2)
  } else x
}
knitr::knit_hooks$set(inline = inline_hook)

```

##Load and Pre-process the Data
Here we need to convert the data into a form suitable for analysis.  I notice that the date column is of
thetype 'factor' and the other two columns are of the type "int".  I convert the date column to a 'date'
class and the other two columns to the "numeric" class.
```{r load_data, echo=TRUE}
data <- read.csv("activity.csv")

#convert to date and numeric types classes.
data$steps <- as.numeric(data$steps)
data$date <- as.Date(data$date)
data$interval <- as.numeric(data$interval)
```

## What is mean total number of steps taken per day?
```{r Mean Steps, echo=TRUE}
cleandata <- data[is.na(data$steps)==FALSE,]
mean_steps <- tapply(cleandata$steps, cleandata$date, sum)
hist(tapply(cleandata$steps, cleandata$date, sum), main = "Histogram of Steps/Day")

AvgSteps <- mean(mean_steps, na.rm=TRUE)
MedSteps <- median(mean_steps, na.rm=TRUE)
```

Mean Total Daily Steps = `r AvgSteps` steps  
Median Total Daily Steps = `r MedSteps` steps
