---
title: "Course 5 Week 2 Project"
output: html_document
---
```{r "setup", include = FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/Christopher/Desktop/Files Not Yet Backed-up/R Files/Course 5/RepData_PeerAssessment1")
```

```{r no sci_notattion}
inline_hook <- function(x) {
  if (is.numeric(x)) {
    format(x, digits = 4)
  } else x
}
knitr::knit_hooks$set(inline = inline_hook)

```

##Load and Pre-process the Data
Here we need to convert the data into a form suitable for analysis.  I notice that the date column is of
thetype 'factor' and the other two columns are of the type "int".  I convert the date column to a 'date'
class and the other two columns to the "numeric" class.
```{r load_data, echo=TRUE}
data <- read.csv("activity.csv")

#convert to date and numeric types classes.
data$steps <- as.numeric(data$steps)
data$date <- as.Date(data$date)
data$interval <- as.numeric(data$interval)
```

## What is mean total number of steps taken per day?
```{r Mean Steps, echo=TRUE}
cleandata <- data[is.na(data$steps)==FALSE,]
mean_steps <- tapply(cleandata$steps, cleandata$date, sum)
hist(tapply(cleandata$steps, cleandata$date, sum), main = "Histogram of Steps/Day", breaks=seq(0,25000,1000))

mean(mean_steps, na.rm=TRUE)
median(mean_steps, na.rm=TRUE)

AvgSteps <- mean(mean_steps, na.rm=TRUE)
MedSteps <- median(mean_steps, na.rm=TRUE)
```

Mean Total Daily Steps = `r AvgSteps` steps  
Median Total Daily Steps = `r MedSteps` steps

## What is the average daily activity pattern?
```{r Avg. Daily activity Pattern, echo=TRUE}

AvgSteps5min <- aggregate( cleandata$steps ~ cleandata$interval, FUN = mean)
names(AvgSteps5min) <- c("Interval","AverageSteps")
head(AvgSteps5min)

plot(AvgSteps5min$Interval, AvgSteps5min$AverageSteps, type = "l", xlab = "5-minute Interval", ylab = "Avg. Steps")
MaxInterval <- AvgSteps5min[which.max(AvgSteps5min$AverageSteps),1]
```

The 5-minute interval with the highest average number of steps is `r MaxInterval` with `r which.max(AvgSteps5min$AverageSteps)` steps.

## Impute Missing Values
```{r CreateImputeCol, echo=TRUE}
names(data) <- c("Steps","Date","Interval")
data$ImputedSteps <- data$Steps
head(data)
```
Notice above that the values of ImputedSteps column are all NA. 

Now we will impute values using the calculated averagesby 5-minute interval.
```{r ImputeValue, echo=TRUE}
for(i in which(is.na(data$Steps))){
    data$ImputedSteps[i] <- AvgSteps5min$AverageSteps[which(AvgSteps5min$Interval == data$Interval[i] )]
}

head(data)
```
Notice now how the values of ImputedSteps column are filled in with average values  
taken from the AvgSteps5min table.

```{r ImputeHisto, echo=TRUE}
hist(tapply(data$ImputedSteps, data$Date, sum), main = "Histogram of Steps/Day", breaks=seq(0,25000,1000))
ImputedAvgSteps <- tapply(data$ImputedSteps, data$Date, sum)

ImputedMean <- mean(ImputedAvgSteps)
ImputedMed <- median(ImputedAvgSteps)
```

The imputed mean number of steps taken per day is `r ImputedMean` steps.  
The imputed median number of steps taken per day is `r ImputedMed` steps.  

Notice that the mean didn't change with the imputed values, but the median change from
`r MedSteps` steps to `r ImputedMed` steps.

## Are there differences in activity patterns between weekdays and weekends?
```{r Weekdays, echo = TRUE}
library(lubridate)
data$Day <- weekdays(as.POSIXct(data$Date))
data$DayNumber <- wday(as.POSIXct(data$Date))
for(i in 1:dim(data)[1]){

if(data$DayNumber[i] == 1){
    data$Weekend[i] <- "Weekend"
} else if(data$DayNumber[i] == 7){
    data$Weekend[i] <- "Weekend"
} else { data$Weekend[i] <- "Weekday"}
}

WkDayAvgSteps5min <- aggregate(ImputedSteps ~ Weekend + Interval, data = data, FUN = mean)

library(lattice)
data <- transform(data, Weekend = factor(Weekend))
xyplot(ImputedSteps ~ Interval | Weekend, data = WkDayAvgSteps5min , layout = c(1,2), type = "l")
```